---
title: "Statistical Inference Course Project-1"
author: "Shivangi"
date: "8/8/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

## PART-1: SIMULATION

In this project you will investigate the exponential 
distribution in R and compare it with the Central Limit 
Theorem. The exponential distribution can be simulated in R 
with rexp(n, lambda) where lambda is the rate parameter. The
mean of exponential distribution is 1/lambda and the 
standard deviation is also 1/lambda. Set lambda = 0.2 for 
all of the simulations. You will investigate the 
distribution of averages of 40 exponentials. Note that you 
will need to do a thousand simulations.


### Observing the population:
In R, we can obtain n samples from an exponential 
distribution with the following code:
```{r}
rexp(10,0.2)
```

Now we take the exponential population with 50,000 subjects.
```{r, results='hide'}
pop <- rexp(50000,0.2)
mu <- mean(pop)
sigma <- sd(pop)
cat("mu= ",mu) 
cat("\nsigma= ",sigma)
```

Histogram of the population having exponential ditribution:

```{r hist}
library(ggplot2)
qplot(pop,col=I("blue"),fill=I("green"),main="Population",binwidth=0.5,xlab = "")

```

### Sampling 40 samples from exponential population:

Distribution of 1000 averages of 40 random exponentials are:

```{r}
mean = NULL
for (i in 1:1000){
mean <- c(mean, mean(rexp(40,0.2)))
}
qplot(mean,col=I("blue"),fill=I("pink"),main="Samples",binwidth=0.1,xlab = "")
```


Although it’s slightly skewed to the right, the bell shape of the Normal distribution is evident. Strikingly, the mean appears to be around 5, which is what we expected.  
  
### Calculating sample mean and standard deviation from distribution function:

Now we calculate the expected mean and variances for the exponential distribution of samples(n) equal to 40.
```{r}
n <-40
lambda <- 0.2
exp_mean <- 1/lambda
exp_sd <- 1/lambda/sqrt(40)
sprintf("expected sample mean = %s",exp_mean)
sprintf("expected sample std deviation = %s",exp_sd)
```



### Comparing the theoretical or expected sample mean of the population from the sample mean:

We know the expected value of sample mean is the estimate of the population mean.  
Mean of sample means is the expected value of mean.

```{r}
cat("The expected mean =",exp_mean)
cat("\t  Actual mean = ",mean(mean))
```
### Comparing the theoretical or estimated sample standard deviated of the population from the sample mean:

It is known that standard deviation of sample mean

```{r}
cat("Expected SD =",exp_sd)
cat("\t Actual SD =",sd(mean))
```
Unsurprisingly, the CLT works.